---
title: "Activity Exploration"
author: "Jasper"
date: "September 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
```

## Getting Data

First off, we have to get the data which is given to us in the repository.
Using `read.csv()` we get a all information in a data frame, and we can use the `lubridate` package to convert the text dates into actual dates.
```{r data import}
    unzip("activity.zip")
    Activities<-read.csv("activity.csv",header=TRUE,na.strings = "NA")
    Activities$date<-ymd(Activities$date)
    CleanedActivities<-filter(Activities,!is.na(steps))
```

The data is now tidy: each row is an observation, and each column is either a value or a description for the observation. It is time to continue the analysis.

## Total number of steps per day

```{r histogram}
    Days<- CleanedActivities%>%
        group_by(date)%>%
        summarize(TotalSteps=sum(steps))
    ggplot(Days,aes(TotalSteps))+geom_histogram(binwidth = 1000)
    # mean <- mean(Days$TotalSteps)
    # median<- median(Days$TotalSteps)
```
The mean number of steps per day is `r mean(Days$TotalSteps)` and the median number is `r median(Days$TotalSteps)`. 

## Daily activity pattern
We use `ggplot` to get the time series for steps per interval.
```{r timeseries average interval}
    AverageDay<-CleanedActivities%>%
        group_by(interval)%>%
        summarize(AverageInterval=mean(steps,na.remove=TRUE))
    ggplot(AverageDay,aes(interval,AverageInterval))+geom_line()
```

##Imputing missing values